---
title: "Customization & Themes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customization & Themes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`dtsmartr` supports deep interface customization through `dtsmartr_options()`. Developers can toggle UI panels, control access to analytical utilities, customize missing cell representations, and switch between light and dark themes.

## 1. Visual Color Themes (Dark Mode)

`dtsmartr` features a modern, premium dark palette designed for clinical environments and night shifts.

* **`theme = "auto"` (Default)**: Automatically detects browser/system settings and matches the theme of your IDE (RStudio or Positron).
* **`theme = "light"`**: Forces the clean light theme.
* **`theme = "dark"`**: Forces the sleek dark theme.

---

## 2. Optional UI Feature Toggles

You can enable or disable individual layout elements depending on the context:

* **`advanced_filter`**: Set `FALSE` to hide the rule-based Query Builder toolbar.
* **`column_picker`**: Set `FALSE` to remove the column dropdown visibility picker in the top-right toolbar.
* **`allow_export`**: Set `FALSE` to remove the clipboard code generator modal button.
* **`header_summary`**: Set `FALSE` to collapse the column headers into clean, flat labels—removing the spark-histograms, top categories bar, and missingness progress indicators.

---

## 3. Custom Missing Value Strings (`na_string`)

Control how missing/null values are represented. Cells containing `NA` or `null` will render as italicized, muted-gray cells:

* **Default**: Displays `"NA"`.
* **Custom**: Set `na_string = "—"` or `na_string = "Missing"` to specify custom text.

---

## 4. Pre-Hidden Columns (`hidden_columns`)

Specify a vector of column names to hide by default on initial rendering. Users can still show them using the column picker (if enabled):

* **Example**: `hidden_columns = c("cyl", "hp")`.

---

## 5. Live Theme & Customization Example

Below is a live interactive grid rendering in **Dark Mode** with pre-hidden variables (`hp`, `vs`), custom missing cell markers (`na_string = "—"`), and the column picker disabled:

```{r}
library(dtsmartr)

dtsmartr(
  data = mtcars,
  options = dtsmartr_options(
    theme           = "dark",
    hidden_columns  = c("hp", "vs"),
    na_string       = "—",
    column_picker   = FALSE
  )
)
```
