---
title: "Filtering & Checklist Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filtering & Checklist Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`dtsmartr` provides powerful column filtering tools, from quick checkboxes to complex logical query rules.

## 1. Column Quick Filters
Click the hamburger icon (≡) on any column to open the dedicated filter panel:
- **Categorical & Low Cardinality Columns**: Columns with 20 or fewer unique values show a checklist. Search values dynamically, check the categories you want, and click **Apply** (or press **Enter**) to filter.
- **Continuous / Numeric Columns**: Shows a text search input box. Type your filter query and press **Enter** to instantly filter matching values.

## 2. Advanced Query Builder
Click the **Advanced Filter** toggle on the top toolbar to reveal the rule-based Query Builder:
- Create nested logical conditions using **AND** or **OR** connector groups.
- Set complex column rules (operators like `equals`, `contains`, `in`, `greater than`, `is null`, etc.).
- Combine quick filters and advanced rule conditions together (they evaluate concurrently).

## 3. Live Interactive Filtering Example
Try searching or filtering columns in the live grid below:

```{r}
library(dtsmartr)
dtsmartr(mtcars)
```
