---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`dtsmartr` is a high-performance, interactive virtualized grid viewer designed to make exploring data in R easy, fast, and reproducible.

## Installation

You can install the development version of `dtsmartr` from GitHub:

```r
# install.packages("devtools")
devtools::install_github("wagh-nikhil/dtsmartr")
```

## Basic Usage

The primary function of the package is `dtsmartr()`. Simply pass any `data.frame` to launch the explorer:

```{r}
library(dtsmartr)

# Explore the classic mtcars dataset
dtsmartr(mtcars)
```

## Lightweight Mode

If you need a clean, lightweight, metadata-free grid viewer for high performance, you can use the newly added `dtsmart_lite()` or its alias `dtsmartr_lite()`:

```{r}
# Forces a compact, clean grid view of a dataset
dtsmart_lite(mtcars, title = "Mtcars Lite")
```

The lightweight mode automatically configures the grid with:
* `advanced_filter = FALSE` (Hides Advanced logical condition rules panel)
* `allow_export = FALSE` (Hides the 'Query Code' button entirely)
* `header_summary = FALSE` (Collapses headers to compact mode, hiding micro-histograms and completeness bars)
