---
title: "Saving HTML Reports & Documents"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Saving HTML Reports & Documents}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`dtsmartr` supports exporting your interactive grid sessions as offline-capable standalone HTML files, as well as embedding them seamlessly inside **R Markdown (.Rmd)** and **Quarto (.qmd)** dynamic reports.

---

## 1. Standalone HTML Reports (`save_dtsmartr`)

Export any data frame as a standalone web app using the `save_dtsmartr()` function:

```r
library(dtsmartr)

# Export as a self-contained portable HTML file
save_dtsmartr(
  data = mtcars, 
  file = "reports/mtcars_dashboard.html"
)
```

The resulting file runs **offline in any browser** without needing R, Shiny, or a network connection.

### Self-Contained vs. Lightweight Output Modes

* **`selfcontained = TRUE` (Default)**: Combines all React code, styling assets, and data base64 streams directly into a single `.html` file. Requires Pandoc to compile.
* **`selfcontained = FALSE`**: Recommended for large datasets or environments without Pandoc. It saves a lightweight `.html` file alongside a `<file>_files/` companion directory containing the shared javascript and style sheets.

---

## 2. R Markdown (.Rmd) Document Integration

You can render `dtsmartr` interactive grids inline in any R Markdown HTML document. 

Simply place the function inside standard code chunks. Set R Markdown chunk options (like `echo = FALSE` or `message = FALSE`) to suppress code display and warnings:

````markdown
---
title: "Clinical Analysis Report"
output: html_document
---

Below is the interactive labs analysis grid:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dtsmartr)
dtsmartr(mtcars)
```
````

When you knit the document, R Markdown compiles the widget as an interactive grid directly inline with your text.

---

## 3. Quarto (.qmd) Document Integration

Quarto supports `dtsmartr` natively out of the box. Add the code chunk using Quarto's standard syntax and comments:

````markdown
---
title: "Study Cohort Review"
format: html
---

## Laboratory Panel Exploration

Explore the interactive subject database:

```{r}
#| echo: false
#| message: false
#| warning: false

library(dtsmartr)
dtsmartr(mtcars)
```
````

When Quarto renders to HTML (`quarto render report.qmd`), the widget embeds inside the page container.

---

## 4. Live Embedded Integration Preview

Here is an actual live rendering of `dtsmartr` embedded inside the R Markdown page container of this article, showing how it looks inside your generated reports:

```{r}
library(dtsmartr)
dtsmartr(mtcars)
```
